---
title: "Assignment_4"
author: "Wilson Ng"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(dplyr)
```

## Getting data from local machine

```{r}
data <- read.csv("/Users/WN/Desktop/flights.csv")

data
```

## I don't have any idea what the numbers represent under each "City" column, but I'm going to move forward with the assumption that they are for arrival delays in minutes.

```{r}
data <- data %>%
  gather("Los.Angeles", "Phoenix", "San.Diego", "San.Fransisco", "Seattle", key="Arrivals", value="Arr_Delay")

data
```

## Dropping rows with NA cells and also dropping rows with empty cells on the Distance column.

```{r}
data <- data %>% drop_na()

data
```

## Filling in the missing cells for the airline column and renaming the first two columns of the dataframe.
```{r}
data <- data %>% filter(data$Arr_Delay != "")

for(row in 1:nrow(data)) {
  if(row %% 2 == 0) {
    data$X[row] = data$X[row - 1]
  }
}

data
```
```{r}
data <- data %>% rename_at('X', ~'Airlines')

data <- data %>% rename_at('X.1', ~'Status')

data$Arr_Delay <- as.numeric(sub(",", "", data$Arr_Delay, fixed = TRUE))

data
```
## Calculating the mean for arr_delay between the two airlines. Judging from the mean - we can conclude that AM WEST has a higher average when it comes to arrival delays in minutes.

```{r calculating the mean for arr_delay between the two airlines}
data %>%
  group_by(Airlines) %>%
  summarise(mean_delay = mean(Arr_Delay))
```